---
title: "PumilioColoromicsWGCNA"
author: "Adam Stuckert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document: default
  pdf_document: default
---

### Setup

```{r setup, include=FALSE}
# load packages required for analyses
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tximport)
library(DESeq2)
library(dplyr)
library(foreach)
library(data.table)
library(splines)
#library(ggthemes)
#library(scales)
library(gridExtra)
#library(tidyr)
#library(pheatmap)
#library(RColorBrewer)
library(ggplot2)
library("BiocParallel")
register(SnowParam(1))
#library(apeglm)
library(tidyverse)
library(WGCNA)
library(flashClust)
```

Print R environment information:

```{R session info}
sessionInfo()
```

This Rmarkdown file is for WGCNA analyses of the Mullerian mimicry RNA seq dataset.

#### WGCNA analyses

```{R data import}
# countdata
adultWGCNAdata <- read.table("data/adultcountdata.tsv", header = TRUE, sep = "\t")
#adultWGCNAdata <- as.integer(adultWGCNAdata)
# sample information
samples <- read.csv("results/samplespreadsheet.csv")
adultWGCNAsamples <- samples %>% filter(age.class == "adult")
# keep only sample name and location
adultWGCNAsamples <- adultWGCNAsamples %>% dplyr::select(sample, location)
# make sample name rowname
adultWGCNAsamples <- adultWGCNAsamples %>% remove_rownames %>% column_to_rownames(var="sample")
#write.csv(samples, "results/samplespreadsheet.csv", row.names = F)
```


```{r}

#################### List all samples from expression data ###############
# get the directory/path for each sample in this study
base_dir <- getwd()
# get directories
dirs <- list.dirs(path = "data/ORP_salmon_quantification", full.names = F, recursive = F)
# append full file path to each sample id
files <- file.path(base_dir, "data/ORP_salmon_quantification", dirs, "quant.sf") # files = directory + salmon directory + sample name + quantifictaion file name
names(files) <- dirs
all(file.exists(files)) # do these all actually exist?

#################### Make sample spreadsheet ###############
samples <- as.data.frame(dirs)
samples$sample <- dirs
samples <- samples[,c(2,1)]

# remove extra bits from sample ID
samples$sample <- gsub(pattern = "_S.*$", replacement = "", samples$sample)

# create age class
samples$age.class <- samples$sample
samples$age.class <- gsub(pattern = "AD.*$", replacement = "adult", samples$age.class)
samples$age.class <- gsub(pattern = "GS.*$", replacement = "tadpole", samples$age.class)

# create a bigger age class one for plotting
samples$full.age.class <- samples$sample
samples$full.age.class <- gsub(pattern = "AD.*$", replacement = "adult", samples$full.age.class)
samples$full.age.class <- gsub(pattern = "_.*$", replacement = "", samples$full.age.class)

# create Gosner stage
samples$gosner.stage <- samples$sample
samples$gosner.stage <- gsub(pattern = "AD.*$", replacement = "", samples$gosner.stage)
samples$gosner.stage <- gsub(pattern = "_.*$", replacement = "", samples$gosner.stage)
samples$gosner.stage <- gsub(pattern = "GS", replacement = "", samples$gosner.stage)

# create sampling locality
samples$location <- samples$sample
samples$location <- gsub(pattern = "^.*_D.*$", replacement = "dorsum", samples$location)
samples$location <- gsub(pattern = "^.*_V.*$", replacement = "venter", samples$location)

adultsamples <- samples %>% filter(age.class == "adult")
rownames(adultsamples) <- adultsamples$sample
adultfiles <- file.path(base_dir, "preliminary_salmon_quantification", adultsamples$dirs, "quant.sf")
names(adultfiles) <- paste0(adultsamples$sample)


#################### Import annotation documents ###############
annos <- fread("annotation_documents/pumilio.annotation.txt", header = FALSE)

# rename columns from annos file
annos <- annos[,c(1,3)]
colnames(annos) <- c("transcript", "Entry")

# import gene names
genes <- fread("annotation_documents/uniprot-reviewed_yes.tab", header = TRUE)

tx2annos <- dplyr::left_join(annos, genes, by = "Entry")
#subset out only tx and protein id columns
tx2gene <- tx2annos[,c(1,3)]
colnames(tx2gene)[2] <- "Entry_name"
tx2gene$Entry_name <- gsub(pattern = "_.*$", replacement = "", tx2gene$Entry_name)


#################### Import expression data ###############
# drop any nas for importing
tx2genenoNA <- drop_na(tx2gene)


#################### Import expression data ###############
# import data via tximport
txiadult <- tximport(adultfiles, type="salmon", tx2gene=tx2genenoNA)
names(txiadult)

adultcounts <- txiadult$counts
write.table(adultcounts, "data/adultcountdata.tsv", sep = "\t")
```

Create DESeq2 object.

```{r DESEQ creation}
# create DESeq dataset
dds <- DESeqDataSetFromTximport(txiadult, colData = adultsamples, design = ~ location)

#dds <- DESeqDataSetFromMatrix(countData = adultWGCNAdata,
         #                              colData = adultWGCNAsamples,
               #                        design = ~ location)

```

```{r, evaluate genes to remove}

## using counts imported via tximports from above
vsd <- vst(dds, blind = FALSE)
vsd <- assay(vsd)
# transpose, as required by WGCNA
vsd <- as.data.frame(t(vsd))

# check if there are gene outliers
gsg = goodSamplesGenes(vsd, verbose = 3)
gsg$allOK

# remove any offenders  
if (!gsg$allOK)
  {if (sum(!gsg$goodGenes)>0)
       printFlush(paste("Removing genes:", paste(names(vsd)[!gsg$goodGenes], collapse= ", ")));
       if (sum(!gsg$goodSamples)>0)
           printFlush(paste("Removing samples:", paste(rownames(vsd)[!gsg$goodSamples], collapse=", ")))
       vsd <- vsd[gsg$goodSamples, gsg$goodGenes]
       }

```

Do they match up?

```{r, eval = F}
# verify data and sample information align
#table(WGCNAsamples$sample==rownames(vsd))
table(adultsamples$sample==rownames(vsd))

# drop lane
#WGCNAsamples <- dplyr::select(adultsamples, -c(filename,lane))
# make sample id row name...
#WGCNAsamples <- data.frame(WGCNAsamples[,-1], row.names=WGCNAsamples[,1])

```

Cluster samples by expression.

```{R WGCNA: cluster by expression}

# first modify sample df to:
adultsamples <-  dplyr::select(adultsamples, -c(sample,dirs,age.class,full.age.class,gosner.stage))


A = adjacency(t(vsd),type="signed") # this calculates the whole network connectivity
k = as.numeric(apply(A,2,sum))-1 # standardized connectivity
Z.k = scale(k)
thresholdZ.k = -2.5 # often -2.5
outlierColor = ifelse(Z.k<thresholdZ.k,"red","black")
sampleTree = flashClust(as.dist(1-A), method = "average")
# Convert traits to a color representation where red indicates high values
# needs to be numeric
adultsamples$location <- as.numeric(as.factor(adultsamples$location))
traitColors = data.frame(numbers2colors(adultsamples,signed=FALSE))
dimnames(traitColors)[[2]] = paste(names(adultsamples))
datColors = data.frame(outlier = outlierColor,traitColors)

plotDendroAndColors(sampleTree,groupLabels=names(datColors),
                    colors=datColors,main="Sample Dendrogram and Trait Heatmap")

```

A possible outlier, but we will leave it be for now.

```{R WGCNA: soft threshold power...}
powers = c(c(1:10), seq(from =10, to=30, by=1)) #choosing a set of soft-thresholding powers
sft = pickSoftThreshold(vsd, powerVector=powers, verbose =5, networkType="signed") #call network topology analysis function

```

```{R WGCNA: scale independence}
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab= "Soft Threshold (power)", ylab="Scale Free Topology Model Fit, signed R^2", type= "n", main= paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels=powers, col="red")
abline(h=0.90, col="red")
```

```{R WGCNA: connectivity}
plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab= "Soft Threshold (power)", ylab="Mean Connectivity", type="n", main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, col="red")
```

Soft power asymptotes pretty hard. I'll attempt this with 4. 

Construct networks.

```{R WGCNA: construct networks}
net = blockwiseModules(vsd, power = 4,
    TOMType = "unsigned", minModuleSize = 30,
    reassignThreshold = 0, mergeCutHeight = 0.25,
    numericLabels = TRUE, pamRespectsDendro = FALSE,
    saveTOMs = TRUE,
    saveTOMFileBase = "AdultpumilioTOM",
    verbose = 3)

```

Plot these modules.

```{R WGCNA: plot modules}
table(net$colors)

# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
    "Module colors",
    dendroLabels = FALSE, hang = 0.03,
    addGuide = TRUE, guideHang = 0.05)
```

Correlate traits.

```{R WGCNA: correlate traits, eval = FALSE}
#Define number of genes and samples
nGenes = ncol(vsd)
nSamples = nrow(vsd)

#Recalculate MEs with color labels
MEs0 = moduleEigengenes(vsd, mergedColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, adultsamples, method = "pearson", use= "all.obs")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
```

Make a figure of this. Heatmap from WGCNA does not work with 1 trait, so don't run for adults.

```{r, adult heatmap, eval = FALSE}
#Print correlation heatmap between modules and traits
textMatrix= paste(signif(moduleTraitCor, 2), "\n(", 
                    signif(moduleTraitPvalue, 1), ")", sep= "")
dim(textMatrix)= dim(moduleTraitCor)
par(mar= c(6, 8.5, 3, 3))
#display the corelation values with a heatmap plot
labeledHeatmap(Matrix= moduleTraitCor, 
            xLabels= colnames(adultsamples), 
            yLabels= names(MEs), 
            ySymbols= names(MEs), 
            colorLabels= FALSE, 
            colors= blueWhiteRed(80), 
            textMatrix= textMatrix, 
            setStdMargins= FALSE, 
            cex.text= 0.5, 
            zlim= c(-1,1), 
            main= paste("Module-trait relationships"))


# save to disk

png("figures/WGCNA_heatmap.png", width = 9.5, height = 7, units = "in", res = 300)
par(mar= c(6, 12, 3, 3))
labeledHeatmap(Matrix= moduleTraitCor, 
            xLabels= names(adultsamples), 
            yLabels= gsub(pattern = "ME", replacement = "", names(MEs)), 
            ySymbols= gsub(pattern = "ME", replacement = "", names(MEs)), 
            colorLabels= FALSE, 
            colors= blueWhiteRed(50), 
            textMatrix= textMatrix, 
            setStdMargins= FALSE, 
            cex.text= 0.5, 
            zlim= c(-1,1), 
            main= paste("Module-trait relationships"))
dev.off()



heatmap(moduleTraitCor,
              name = "Cor",
              column_title="Module-trait relationships",
              rect_gp = gpar(col = "black"),
              cell_fun = function(j, i, x, y, width, height, fill) {
              grid.text(sprintf(textMatrix[i, j]), x, y, gp = gpar(fontsize = 10))})



```

Get module membership. ######## have p values up to here....

```{r WGCNA: output module membership}

datME=moduleEigengenes(vsd,mergedColors)$eigengenes
datKME=signedKME(vsd, datME, outputColumnName="MM.")
genes=names(vsd)
geneInfo0 = data.frame(gene=genes,moduleColor=mergedColors)
color=data.frame(geneInfo0,datKME) #these are from your original WGCNA analysis 
head(color)
write.csv(as.data.frame(color), file = "results/GeneModule_membership.csv")

#### MM pvalues
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(vsd, MEs, use = "p"));
MMPvalue=as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
pvals=data.frame(geneModuleMembership,MMPvalue)
head(pvals)
write.csv(as.data.frame(pvals), file = "results/GeneModule_membership_pvalues.csv")

```

Gene significance by module membership. 

```{R WGCNA: significance of population membership, eval = FALSE}

whichTrait="location" #Replace this with the trait of interest

location = as.data.frame(adultsamples$location);
names(location) = "location"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(vsd, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(vsd, location, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(location), sep="");
names(GSPvalue) = paste("p.GS.", names(location), sep="");

```

Significant gene modules by location

```{R, WGNCA: identify significant modules by location}
# subets out only significant modules
modulePs <- as.data.frame(moduleTraitPvalue)
sigmods <- rownames(subset(modulePs, location < 0.05))
print(paste0("Number of significant modules by color (location): ", length(sigmods)))

```
Create a merged table of modules that are stastically significant for results section.

```{R table of statistically significant modules}
CorVals <- as.data.frame(moduleTraitCor)
PVals <- as.data.frame(moduleTraitPvalue)

ModuleStats <- cbind(CorVals, PVals)
colnames(ModuleStats) <- c("Correlation", "pvalue")

# subset by only significant modules
ModuleStats <- ModuleStats %>% filter(pvalue < 0.05)

write.csv(ModuleStats, "results/WGCNAresults/ModuleStatistics.csv", row.names = TRUE)
```

### Rank genes from significant modules for GO analyses

Get gene information for these.

```{R, WGNCA: gene information for significant modules by morph}
MMsigmods <- gsub(pattern = "ME", replacement = "MM.", sigmods)
#setDF(datKME)
#datKME %>% dplyr::select(datKME, matches(paste(MMsigmods, collapse="|")))
#datKME[,c(sigmods)]

#sigmorphmodules <- datKME[,c("MM.pink", "MM.purple", "MM.grey")]


# loop to make ordered outputs for GOrilla
for (i in 1:length(MMsigmods)){
  modcol <- MMsigmods[i]
  # get data for module color
  tmp <- datKME[,modcol]
  gene_rows <- rownames(datKME)
  tmp <- as.data.frame(cbind(gene_rows, as.numeric(tmp)))
  colnames(tmp) <- c("gene", "module.membership")
  tmp$module.membership <- as.numeric(as.character(tmp$module.membership))
  tmp <- tmp[order(tmp[2], decreasing = TRUE), ]
  write.csv(tmp, paste0("results/WGCNAresults/adultlocation_", modcol, "ModuleMembership.csv"), row.names = FALSE)
}
```







############ this is all good until here! I may or may not include any of the stuff below. If it is interesting at at all....





## What are my hub genes?

```{R Hub genes}

# remove "ME" from all sigmods
sigmods <- gsub(pattern = "ME", replacement = "", sigmods)
probes = names(vsd)
nTop = 30
moduleColors=mergedColors
adj = adjacency(vsd, power = 10, type = "signed") #specify network type

## Translate adjacency into topological overlap matrix and calculate the corresponding dissimilarity

TOM = TOMsimilarity(adj, TOMType="signed") #Topological Overlap Matrix
dissTOM = 1-TOM 

for(i in 1:length(unique(sigmods))){
  #inModule = sigmods[i];
  #modProbes = probes[inModule];
  inModule = (moduleColors == sigmods[i]);
  modProbes = probes[inModule];
  ## Select the corresponding Topological Overlap
  
  modTOM = TOM[inModule, inModule];
  dimnames(modTOM) = list(modProbes, modProbes)
  
  ## Genes with highest connectivity within each significant module
  
  IMConn = softConnectivity(vsd[, modProbes])
  top = (rank(-IMConn) <= nTop)
  
  tempdat<-data.frame(IMConn,modProbes)[order(-IMConn),]
  names(tempdat)<-c("IMConnectivity","target_id")
  tempdat$ModuleColor <- rep(sigmods[i],nrow(tempdat))
  #tempdat2 = left_join(tempdat,preGO1, by = "target_id")
  write.csv(tempdat, paste0("results/WGCNAresults/HubGenes",sigmods[i],".csv", sep = "_"), na="", row.names=FALSE)
}

## List all the files

temp <- list.files(path = "results/WGCNAresults", pattern = "*HubGenes")
temp <- paste("results/WGCNAresults/", temp, sep = "")
Hub_genes = do.call(rbind, lapply(temp, function(x) read.csv(x, stringsAsFactors = FALSE)))
write.csv(Hub_genes, paste("results/WGCNAresults/HubGenes_all.csv",sep="_"), na="", row.names=FALSE)

```

```{R HUBGENES2}

# Recalculate topological overlap if needed
#TOM = TOMsimilarityFromExpr(vsd, power = 10)

# Select modules
sigmods # Lists modules of interest
module = "yellow"    
moduleColors = mergedColors
# Select module probes
probes = names(vsd)
inModule = is.finite(match(moduleColors, module));
modProbes = probes[inModule];
#modGenes = annote_df$gene_name[match(modProbes, annote_df$target_id)];

# Select the corresponding Topological Overlap
modTOM = TOM[inModule, inModule];
dimnames(modTOM) = list(modProbes, modProbes)
cyt = exportNetworkToCytoscape(modTOM,
       edgeFile = paste("results/WGCNAresults/CytoscapeInput-edges-", paste(module, collapse="-"), ".txt"),
       nodeFile = paste("results/WGCNAresults/CytoscapeInput-nodes-", paste(module, collapse="-"), ".txt"),
       weighted = TRUE,
       threshold = 0.115, # Change to get about 30
       nodeNames = modProbes,
       #altNodeNames = modGenes,
       nodeAttr = moduleColors[inModule]);
cyt

## Record of thresholds:
# Dark red - 0.056
# black - 0.25
# red - 0.145
# paleturquoise - 0.045
# salmon - 0.06
# cyan - 0.08
# midnightblue - 0.07
# yellow - 0.115


cytoscapePing()

```

